SunPy - Python for Solar Physics

### 2014-01-31 2-Data_Types.tex

SunPy currently provides data strucutures that are specifically designed for the
three primary varieties of solar physics data: images, time series, and
spectra. These core data types are supported by SunPy's classes:
\texttt{Map} (2D spatial data), \texttt{LightCurve} (1D temporal series),
and \texttt{Spectrum} and \texttt{Spectrogram} (1D and 2D spectra).

@@

This section will give a brief overview of the \textit{current} functionality 
of each of these data types.

\subsection{Visualisation}
\label{subsec:Viz}
All of the core SunPy data types 
include visualisation methods that are tailored to each data type. 
These visualisation methods all currently utilise the \texttt{matplotlib} 
package and are designed in such a way that they integrate well with 
the \texttt{pyplot} functional interface of \texttt{matplotlib}.

@@

>>>>

The core of SunPy are a set of data strucutures that are specifically designed for the
three primary varieties of solar physics data: images, time series, and
spectra. These core data types are supported by SunPy's classes:
\texttt{Map} (2D spatial data), \texttt{LightCurve} (1D temporal series),
and \texttt{Spectrum} and \texttt{Spectrogram} (1D and 2D spectra).

@@

All of the core SunPy data types 
include visualisation methods that are tailored to each data type. 
These visualisation methods all currently utilise the \texttt{matplotlib} 
package and are designed in such a way that they integrate well with 
the \texttt{pyplot} functional interface of \texttt{matplotlib}.

@@

This section will give a brief overview of the \textit{current} functionality 
of each of the core SunPy data types.

### 2014-01-31 3-4-Helioviewer.tex

SunPy provides the ability to download images hosted by the
Helioviewer Project (\href{http://helioviewer.org}{http://helioviewer.org}).  
The aim of the Helioviewer Project is to enable
the exploration of solar and heliospheric data from multiple data
sources (such as instrumentation and feature/event catalogues) via
easy-to-use visual interfaces. The Helioviewer Project have developed two client 
applications that allow users to browse images and create movies of the Sun taken 
by a variety of instruments: \href{http://www.helioviewer.org}{Helioviewer.org}, a 
Google Maps-like web application, and \href{http://www.jhelioviewer.org}{JHelioviewer}, 
a movie streaming desktop application. In order to manage all of this data, the Helioviewer
project maintains archives of all of its data in JPEG2000 format (for further
details on the Helioviewer Project see \cite{muller2009}). The
JPEG2000 files are typically highly compressed compared to the source
FITS files from which they are generated, but are still high-fidelity, and thus can be used to quickly
visualise large amounts of data from multiple sources.  SunPy is
also used in Helioviewer production servers to manage the download and
ingestion of JPEG2000 files from remote servers.

>>>>

SunPy provides the ability to download images hosted by the
Helioviewer Project (\href{http://helioviewer.org}{http://helioviewer.org}).  
The aim of the Helioviewer Project is to enable
the exploration of solar and heliospheric data from multiple data
sources (such as instrumentation and feature/event catalogues) via
easy-to-use visual interfaces. The Helioviewer Project have developed two client 
applications that allow users to browse images and create movies of the Sun taken 
by a variety of instruments: \href{http://www.helioviewer.org}{Helioviewer.org}, a 
Google Maps-like web application, and \href{http://www.jhelioviewer.org}{JHelioviewer}, 
a movie streaming desktop application. In order to manage all of this data, the Helioviewer
project maintains archives of all of its data in JPEG2000 format (for further
details on the Helioviewer Project see \citealt{muller2009}). The
JPEG2000 files are typically highly compressed compared to the source
FITS files from which they are generated, but are still high-fidelity, and thus can be used to quickly
visualise large amounts of data from multiple sources.  SunPy is
also used in Helioviewer production servers to manage the download and
ingestion of JPEG2000 files from remote servers.

### 2014-01-31 6-Future.tex

SunPy as a project has existed for roughly three years. In this time, 
the code base has grown to over 17,000 lines. SunPy is already a 
useful package for the analysis of calibrated solar data, and it
continues to gain significant new capabilities with each successive release.
The primary focus of the 
SunPy library is the analysis and visualisation of `high-level' solar 
data. This means data that has been put through instrument processing 
and 
calibration routines, and contains full (WCS) coordinate information. 
The plan for SunPy is to continue development within this 
scope. The 
primary components of this plan are to provide a set of data types 
that are 
interchangeable with one another: e.g., if you slice a 
\texttt{MapCube} 
along one spatial coordinate, a \texttt{LightCurve} of intensity along the 
time range of 
the \texttt{MapCube} should be returned. To achieve this goal, all the 
data 
types need to share a unified coordinate system architecture so that 
each data 
type is aware of what the physical type of its data is and how 
operations on 
that data should be performed. This will enable useful operations
such as the coordinate and solar-rotation-aware 
overplotting of HELIO (Section \ref{ssec:helio}) and HEK
results (Section \ref{ssec:hek}) onto maps (Section \ref{ssec:map}).

@@

In the future, additional HELIO services will be supported by SunPy, aiming for
seamless interaction between the other services and tools available (e.g., 
\texttt{hek}, \texttt{map}).

>>>>

Over the three years of SunPy's development, the code base has grown to over 17,000 lines. 
SunPy is already a 
useful package for the analysis of calibrated solar data, and it
continues to gain significant new capabilities with each successive release.
The primary focus of the 
SunPy library is the analysis and visualisation of `high-level' solar 
data. This means data that has been put through instrument processing 
and 
calibration routines, and contains full (WCS) coordinate information. 
The plan for SunPy is to continue development within this 
scope. The 
primary components of this plan are to provide a set of data types 
that are 
interchangeable with one another: e.g., if you slice a 
\texttt{MapCube} 
along one spatial coordinate, a \texttt{LightCurve} of intensity along the 
time range of 
the \texttt{MapCube} should be returned. To achieve this goal, all the 
data 
types need to share a unified coordinate system architecture so that 
each data 
type is aware of what the physical type of its data is and how 
operations on 
that data should be performed. This will enable useful operations
such as the coordinate and solar-rotation-aware 
overplotting of HELIO (Section \ref{ssec:helio}) and HEK
results (Section \ref{ssec:hek}) onto maps (Section \ref{ssec:map}).
Finally, support for new data providers and services will be integrated into SunPy.
For example, new HELIO services will be supported by SunPy, aiming for
seamless interaction between the other services and tools available (e.g., 
\texttt{hek}, \texttt{map}).

@@

### 2014-01-31 Summary.tex

\section{Summary}
We have presented the release of SunPy (v0.4), a Python package for solar physics. In
this paper we have described the main functionality which includes the SunPy data types, 
\texttt{Map} (see Section~\ref{ssec:map}), \texttt{Lightcurve} (see Section~\ref{ssec:lightcurve}), and \texttt{Spectrogram} (see Section~\ref{sec:spectra}).
We have described the data and event catalogue retrieval capabilities of SunPy for
the Virtual Solar Observatory (see Section~\ref{ssec:vso}), the Heliophysics Event Knowledgebase (see Section~\ref{ssec:hek}), as well as
the Heliophysics Integrated Observatory (see Section~\ref{ssec:helio}). We described
a new organization tool for data files integrated into SunPy (see Section~\ref{ssec:db}) 
and we discussed the community aspects, development model (see Section~\ref{sec:dev}), and future plans (see Section~\ref{sec:future}) for the project.
We invite members of community to contribute to the effort by using SunPy for their 
research, reporting bugs, and sharing new functionality with the project.

>>>>

\section{Summary}
We have presented the release of SunPy (v0.4), a Python package for solar physics. In
this paper we have described the main functionality which includes the SunPy data types, 
\texttt{Map} (see Section~\ref{ssec:map}), \texttt{Lightcurve} (see Section~\ref{ssec:lightcurve}), and \texttt{Spectrogram} (see Section~\ref{sec:spectra}).
We have described the data and event catalogue retrieval capabilities of SunPy for
the Virtual Solar Observatory (see Section~\ref{ssec:vso}), the Heliophysics Event Knowledgebase (see Section~\ref{ssec:hek}), as well as
the Heliophysics Integrated Observatory (see Section~\ref{ssec:helio}). We described
a new organization tool for data files integrated into SunPy (see Section~\ref{ssec:db}) 
and we discussed the community aspects, development model (see Section~\ref{sec:dev}), and future plans (see Section~\ref{sec:future}) for the project.
We invite members of the community to contribute to the effort by using SunPy for their 
research, reporting bugs, and sharing new functionality with the project.
